An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2017-03-16 uai2017/paper.tex
Slight modifications to logistic regression.
I had mis-run the earlier trial by not repeating samples if they were rejected. That's all taken care of now and the paper has been updated along with figures.

% Daniel: moving this here so figures don't overflow to next pages. We can tune
% the positioning if needed.
\begin{figure*}[t]
	\centering
	\includegraphics[width=1.0\linewidth]{logistic_regression_error_regions.png}
	\caption{
    Binary classification accuracy of the MCMC methods on the 1s vs 7s logistic
    regression task for MNIST-13k (left plot) and MNIST-100k (middle plot) as a
    function of cumulative data usage.  The right plot reports performance of
    {\sc MHminibatch} on both datasets when the temperature starts at 100 and
    drops to 1 after a ``burn-in'' period of 200 samples (vertical dashed line)
    of $\theta$.  For all three plots, one standard deviation is indicated by
    the shaded error regions.
    }
	\label{fig:logistic_performance}
\end{figure*}

@@

The first two subplots of Figure~\ref{fig:logistic_performance} display the
prediction accuracy on both datasets for all methods as a function of the
cumulative training points processed.\footnote{The curves do not span the same
length over the x-axis since the methods consume different amounts of data.} To
generate the curves, for each of the sampled vectors $\theta_t$,
$t\in\{1,\ldots,T\}$, we use $\theta_t$ as the logistic regression parameter.
The results indicate that our test is more efficient, obtaining convergence more
than an order of magnitude faster than {\sc AustereMH(nc)} and several orders of
magnitude compared to {\sc AustereMH(c)} and {\sc MHSubLhd}.  We also observe
the advantage of having higher temperature from the third plot in
Figure~\ref{fig:logistic_performance}. During the exploration period, the
accuracy rapidly increases, and then after 200 samples, we switch the
temperature to 1, but this requires the step size to decrease, hence the smaller
changes in accuracy.

Figure~\ref{fig:logistic_histograms_mnist8m} shows log-log histograms of
minibatch sizes for the three methods on MNIST-100k.
(Figure~\ref{fig:logistic_histograms_mnist} in Appendix~\ref{app:logistic}
contains results for MNIST-13k.) The histograms only represent one
representative trial; Table~\ref{tab:logistic} contains the average of the
average minibatch sizes ($\pm$ one standard deviation) across all trials. {\sc
MHminibatch}, with average minibatch sizes of roughly 129.8 and 202.5 for
MNIST-13k and MNIST-100k, respectively, consumes more than 5x fewer data points
than the next-best method, {\sc AustereMH(nc)}.  We reiterate, however, that
both {\sc AustereMH(nc)} and {\sc MHSubLhd} require computing $\log
p(x_i|\theta)$ and $\log p(x_i|\theta')$ for all $x_i$ each iteration. Our
results here do not count that extra data consumption. Only our method and {\sc
AustereMH(c)} rely solely on the minibatch of data each iteration.

@@

\begin{table}[t]
\caption{Average minibatch sizes ($\pm$ one standard deviation) on logistic
regression on MNIST-13k and MNIST-100k.  The averages are taken over 10
independent trials (5000 samples each) for MNIST-13k and 5 independent trials
(3000 samples each) for MNIST-100k.}
\small % Daniel: yeah, sorry, but otherwise it doesn't fit in a column
\label{tab:logistic}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{Method/Data} & \textbf{MNIST-13k} & \textbf{MNIST-100k}  \\
\hline \\
{\sc MHminibatch}   & $129.8\pm 3.4$    & $202.5 \pm 13.6$ \\
{\sc AustereMH(nc)} & $973.8\pm 49.8$   & $1098.3 \pm 44.9$ \\
{\sc AustereMH(c)}  & $1924.3\pm 52.4$  & $2795.6 \pm 364.0$ \\
{\sc MHSubLhd}      & $10783.4\pm 78.9$ & $14977.3 \pm 582.0$ \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}

>>>>

% Daniel: moving this here so figures don't overflow to next pages. We can tune
% the positioning if needed.
\begin{figure*}[t]
	\centering
	\includegraphics[width=1.0\linewidth]{logistic_regression_error_regions.png}
	\caption{
    Binary classification accuracy of the MCMC methods on the 1s vs 7s logistic
    regression task for MNIST-13k (left plot) and MNIST-100k (middle plot) as a
    function of cumulative data usage.  The right plot reports performance of
    {\sc MHminibatch} on both datasets when the temperature starts at 100 and
    drops to 1 after a ``burn-in'' period of 400 samples (vertical dashed line)
    of $\theta$.  For all three plots, one standard deviation is indicated by
    the shaded error regions.
    }
	\label{fig:logistic_performance}
\end{figure*}

@@

The first two subplots of Figure~\ref{fig:logistic_performance} display the
prediction accuracy on both datasets for all methods as a function of the
cumulative training points processed.\footnote{The curves do not span the same
length over the x-axis since the methods consume different amounts of data.} To
generate the curves, for each of the sampled vectors $\theta_t$,
$t\in\{1,\ldots,T\}$, we use $\theta_t$ as the logistic regression parameter.
The results indicate that our test is more efficient, obtaining convergence more
than an order of magnitude faster than {\sc AustereMH(nc)} and several orders of
magnitude compared to {\sc AustereMH(c)} and {\sc MHSubLhd}.  We also observe
the advantage of having higher temperature from the third plot in
Figure~\ref{fig:logistic_performance}, which plots average performance and one
standard deviation for {\sc MHminibatch} over 10 trials. During the exploration
period, the accuracy rapidly increases, and then after 400 samples, we switch
the temperature to 1, but this requires the step size to decrease, hence the
smaller changes in accuracy.

Figure~\ref{fig:logistic_histograms_mnist8m} shows log-log histograms of
minibatch sizes for the methods on MNIST-100k.
(Figure~\ref{fig:logistic_histograms_mnist} in Appendix~\ref{app:logistic}
contains results for MNIST-13k.) The histograms only represent one
representative trial; Table~\ref{tab:logistic} contains the average of the
average minibatch sizes ($\pm$ one standard deviation) across all trials. {\sc
MHminibatch}, with average minibatch sizes of 133.7 and 242.4 for MNIST-13k and
MNIST-100k, respectively, consumes more than 7x and 4x fewer data points than
the next-best method, {\sc AustereMH(nc)}.  We reiterate, however, that both
{\sc AustereMH(nc)} and {\sc MHSubLhd} require computing $\log p(x_i|\theta)$
and $\log p(x_i|\theta')$ for all $x_i$ each iteration. Our results here do not
count that extra data consumption. Only our method and {\sc AustereMH(c)} rely
solely on the minibatch of data each iteration.

@@

\begin{table}[t]
\caption{Average minibatch sizes ($\pm$ one standard deviation) on logistic
regression on MNIST-13k and MNIST-100k.  The averages are taken over 10
independent trials (5000 samples each) for MNIST-13k and 5 independent trials
(3000 samples each) for MNIST-100k.}
\small % Daniel: yeah, sorry, but otherwise it doesn't fit in a column
\label{tab:logistic}
\vskip 0.15in
\begin{center}
\begin{tabular}{l l l l}
\textbf{Method/Data} & \textbf{MNIST-13k} & \textbf{MNIST-100k}  \\
\hline \\
{\sc MHminibatch}   & $133.7\pm 4.7$    & $242.4 \pm 12.4$ \\
{\sc AustereMH(nc)} & $973.8\pm 49.8$   & $1098.3 \pm 44.9$ \\
{\sc AustereMH(c)}  & $1924.3\pm 52.4$  & $2795.6 \pm 364.0$ \\
{\sc MHSubLhd}      & $10783.4\pm 78.9$ & $14977.3 \pm 582.0$ \\
\end{tabular}
\end{center}
\vskip -0.1in
\end{table}
